%!TEX root = ../main.tex
% Chapter Template



\chapter{Introduction} % Main chapter title
\label{ch:Introduction} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------\section{Personal Statement}
% Alice: "I've implemented a system that tolerates the presence of any attacker"\\
% % Bob: "Which attacker?"\\
% % Alice: "A trivial attacker; one that does nothing."\\
% Bob: "What does your system do?"\\
% Alice: "Nothing"\\
% Bob: "Oh! That I can believe."\\

% \todo[inline]{TOO GENERAL}
% A \emph{secure system} preserves its integrity, confidentiality and availability in the presence of adversaries~\cite{CIAModel}. 
% %A \emph{robust system} preserves its integrity in the presence of adversaries.
% Not all adversaries are the same; a passive attacker only observes, and while this attacker may not compromise the integrity or the availability of the system, they may compromise its confidentiality. 

% %What does it mean for a system to work in the presence of an adversary?
% \begin{itemize}
%     \item There have been not many uses for coalgebra and security.
%     \item Noninterference is a behavioural property 
% \end{itemize}

% Modern systems must not only be correct, but also secure. Proving that a system is correct is a difficult problem, but I would say that proving that a system is secure is exponentially more difficult. Correctness proofs study properties of the system in isolation or under strong interaction guarantees, i.e., the environment never misbehaves. 

Basin and Cremers~\cite{KnowYourEnemy} introduced \emph{Scyther}~\cite{Scyther}, a protocol analysis tool that classifies security protocols in a hierarchy given a set of attacker models. 
The position of a protocol in a hierarchy depends on its level of assurance with respect to a state property: protocols higher in the hierarchy can tolerate the presence of more powerful attackers, i.e., attackers with more capabilities. More precisely, the protocol hierarchy is directed acyclic graph whose nodes are pairs of a protocol and a set of attacker capabilities, where two nodes share an edge if the source protocol-set of capabilities satisfies the state property, but adding an additional capabilities would break the state property in the protocol, whilst the property is still satisfied in the destination node despite the addition of those capabilities.

To check if a protocol satisfies a state property (e.g. perfect forward secrecy) in the presence of adversaries, Basin and Cremers model the protocol as symbolic transition system, and they model an adversary as a set of capabilities that enable previously inexistent transitions in the protocol. These new transitions expand the set of reachable states in the protocol, and this new set of reachable states may contain a state where the knowledge of the attacker has enough sensitive information to break the state property being verified . 

The idea of an attacker adding new transitions to an existing system is simple yet powerful. However, to which extend can an attacker add new transitions to a system? Can they add arbitrary transitions? Can they also remove existing transitions? All these are pertinent questions, and their answers have interesting implications. Basin and Cremers~\cite{KnowYourEnemy} do not allow attackers to create arbitrary transitions; attackers can only create them based on the following three aspects: \emph{which} kind of data is compromised, \emph{whose} data is it, and \emph{when} the compromise occurs. Arbitrary transitions are theoretically possible, but they may be practically infeasible; for example, we could create a transition from a state where the attacker has no knowledge about the secrets to a state where the attacker has knowledge of all of the secrets, and such attacker could break any confidentiality protocol, but the realism of such attacker model is questionable. 

We face a difficult challenge: how can we define realistic attacker models for transition systems? 

%If an attacker can arbitrarily change the transition structure of a system, is it possible to defend against such attacker? 
%It is an idea compatible with both model checking and testing, which is how we can obtain hierarchies of systems. 
%While Basin and Cremers study the effect of attackers for confidentiality, we are interested in studying the effect of attackers for integrity. Clarkson and Schneider write~\cite{QuantitativeIntegrity} 


%  span over three dimensions: \emph{which} data is compromised, \emph{whose} data is it, and \emph{when} the compromise occurs. In a nutshell, a they model a protocol 

% %Attackers that are powerful can replicate the actions of weaker attackers, 

% Pairing systems and attacker models to obtain a position



% In particular, \emph{cyber-physical systems} (CPS), which comprise a majority of the critical infrastructure in many countries, 

% \emph{Universal coalgebra}~\cite{UniversalCoalgebra} 





% %\todo[inline]{Attacker models are usually informal,}
% % ``Can we attack a Turing Machine?" It seems a pointless question; Turing machines are models of computation, not real machines, so why bother pondering about this apparently nonsensical question? 

% Alice: ``I have a proof that my system is better than yours, Bob."\\
% Bob: ``Seems unlikely, show me.''\\
% Alice brings out a piece of paper, and points to two numbers, one clearly bigger than the other. "I ran an analysis on both systems, and mine scored better than yours" says Alice.






\todo[inline]{Biba Duality}
\todo[inline]{Basin Cremers attackers as sets of effects KYE}
\todo[inline]{Confidentiality: test if attackers have some knowledge; Integrity: test if the system is too corrupted.}
\todo[inline]{What, when, how? Modelling attacks}
\todo[inline]{if we know what attacker models are mathematically, we can ask the computer to do analysis for us in integrity. }
\todo[inline]{YOU HAVE TO ADDRESS the problem for confidentiality and cite relevant papers, and quantificaiton}
\todo[inline]{What does Unifying Facets of Information Integrity say?}

\todo[inline]{The idea is that there is a lot of stuff, but they ultimately belong to a bigger set of ``program correctness''. We argue that those are behavioural properties}

\todo[inline]{Coalgebras provide a clear view of the set of behavioural properties for families of systems. That makes them really attractive for studying. Not only that, they offer generalised processes for the incorporation of computational effects (e.g., non-determinism, exceptions, termination, etc.) because of category theory.}

Probably the most important is to motivate the quantification of attackers and how it gave rise to the general treatment of latent behaviours.
\todo[inline]{}
\todo[inline]{The first thing we need to do is describe why latent behaviours are interesting. This requires a historical context. }
\section{Coalgebras}
\todo[inline]{This section explains the importance of coalgebras as a modelling system, why they are so compatible with computability (monads), and in general why they are useful.}
\todo[inline]{I could probably be inspired on how to structure this section by publications on coalgebras. There is a large body of literature I can use.}
\todo[inline]{At least cover that systems are of a form $\delta\colon X\rightarrow D$ where $D$ is some domain, and that there is a way to go from $D$ back to $X$ to continue computation.}
\todo[inline]{Final coalgebras are important because $D=X$, so it is obvious how to continue.}

\section{Latent Behaviours}
\todo[inline]{Explain that in this document we explore the notion of creating new behaviours through transformations of the state space of systems.}
\subsection{Intuition}
\todo[inline]{We present a light version of latent behaviours: you normally start at $X$, then you go to $D$ and then back to $X$, but what if we transformed $X$?}
In latent behaviours we use a transformation $X\rightarrow X$ to affect the behaviour of the system. This transformation is an abstraction; the original system naturally transforms to a system $X\rightarrow D \rightarrow X$ using composition, also of type $X\rightarrow X$. Each function $X\rightarrow X$ can be seen as a black-box of some behaviour. Their interweaving means that they are acting simultaneously. 

\section{Research Questions}
We consider the following broad yet interesting research questions that are pervasive in systems security to motivate our work.
\begin{question}
\label{que:AttackerModel}
How do we precisely describe and generate attacker models, attacks and attackers?
\end{question}
\begin{question}
\label{que:Quantification}
How do we efficiently quantify the effect that a given attacker has on a given system? 
\end{question}
\begin{question}
\label{que:Classification}
How do we compare attacker models and attackers with respect to the effect they potentially have on a given system?
\end{question}
% \begin{question}
% \label{que:Countering}
% How can we counter or mitigate the effect that an attacker has on the system?
% \end{question}
\begin{question}
\label{que:Repair}
How do we repair a system that is vulnerable with respect to a given attacker model?
\end{question}
We do not pretend to solve these questions at this broad level, but we address concrete instances of these questions, which appear as we instantiate systems, attacker models, security requirements, etc. in the following sections.
\section{Applications}
\subsection{Attacker Classification}
How can we model attackers? How can we model their actions? How can we prepare against them?
\subsection{Cyber-Physical System Redesign}
Many attacks are discovered through testing: an experienced hacker creates a proof of concept to illustrate how some system may be attacked. Then, an abstraction is created to explain why the attack works, and to suggest countermeasures against it. 
\subsection{Program Repair}
%Fuzzing is probably one of the most popular techniques for system transformations. 

% \todo[inline]{Actually, the thesis would be super interesting if potential behaviour problems are presented in terms of a game between the system and the attacker, with the attacker following some sort of reactive strategy, and the system as well. At some point, one or the other wins because it is a deterministic game if the system is deterministic. }



% \todo[inline]{You can probably show several papers that display this pattern.} 


% %-----------------------------------
% %	SUBSECTION 1
% %-----------------------------------
% \subsection{Subsection 1}

% Nunc posuere quam at lectus tristique eu ultrices augue venenatis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Aliquam erat volutpat. Vivamus sodales tortor eget quam adipiscing in vulputate ante ullamcorper. Sed eros ante, lacinia et sollicitudin et, aliquam sit amet augue. In hac habitasse platea dictumst.

% %-----------------------------------
% %	SUBSECTION 2
% %-----------------------------------

% \subsection{Subsection 2}
% Morbi rutrum odio eget arcu adipiscing sodales. Aenean et purus a est pulvinar pellentesque. Cras in elit neque, quis varius elit. Phasellus fringilla, nibh eu tempus venenatis, dolor elit posuere quam, quis adipiscing urna leo nec orci. Sed nec nulla auctor odio aliquet consequat. Ut nec nulla in ante ullamcorper aliquam at sed dolor. Phasellus fermentum magna in augue gravida cursus. Cras sed pretium lorem. Pellentesque eget ornare odio. Proin accumsan, massa viverra cursus pharetra, ipsum nisi lobortis velit, a malesuada dolor lorem eu neque.

% %----------------------------------------------------------------------------------------
% %	SECTION 2
% %----------------------------------------------------------------------------------------

% \section{Main Section 2}

% Sed ullamcorper quam eu nisl interdum at interdum enim egestas. Aliquam placerat justo sed lectus lobortis ut porta nisl porttitor. Vestibulum mi dolor, lacinia molestie gravida at, tempus vitae ligula. Donec eget quam sapien, in viverra eros. Donec pellentesque justo a massa fringilla non vestibulum metus vestibulum. Vestibulum in orci quis felis tempor lacinia. Vivamus ornare ultrices facilisis. Ut hendrerit volutpat vulputate. Morbi condimentum venenatis augue, id porta ipsum vulputate in. Curabitur luctus tempus justo. Vestibulum risus lectus, adipiscing nec condimentum quis, condimentum nec nisl. Aliquam dictum sagittis velit sed iaculis. Morbi tristique augue sit amet nulla pulvinar id facilisis ligula mollis. Nam elit libero, tincidunt ut aliquam at, molestie in quam. Aenean rhoncus vehicula hendrerit.